---
title: "**【Data Processing】**"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{css echo = FALSE}
/*-- scss:defaults --*/

body{
  color:#7CA8BF;
  background-color:#ffe4e1;
  font-size: 16px;
}

a{
  color: #ffa9d8;
}

/* width */
::-webkit-scrollbar {
    width: 10px;
    background-color:rgba(255, 192, 203, 0.183)
}

/* Track */
::-webkit-scrollbar-track {
    border-radius: 10px;
}

/* Handle */
::-webkit-scrollbar-thumb {
    border-radius: 10px;
    background-color: #7CA8BF;
}
/* Handle on hover */
::-webkit-scrollbar-thumb:hover {
    width: 40px;
    background-color: #fbcac4;
}

hr{
  border-top: 1px solid #bdcfda;
}

caption {
  display: table-caption;
  text-align: center;
  font-size: 16px;
}

.dataframe{
    width: 100%;
    height: 130px;
    border: 5px solid #bdcfda;
    border-collapse: collapse;
    font-size: 18px;
}

.dataframe tr, .dataframe tr td{
    border: 2px solid #bdcfda;
    align-items: center;
}

.dataframe tr th{
    border: 2px solid #bdcfda;
}

tbody{
  border: 5px solid #bdcfda;
}

.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #ffe4e1;
    background-color: #7CA8BF;
    border-color: #7CA8BF;
}

.code-output{
  color:#ffe4e1;
  background-color:#7CA8BF;
}

```

****

## **0 Data**

```{r message=FALSE, warning=FALSE}
df <- read.csv("csv/20240805基本資料_retrospective.csv",
               na.strings = c("", "NA"),
               fileEncoding = "Big5")

df
```

### 0.1 variables

```{r message=FALSE, warning=FALSE}
summary(df)
```

### 0.2 variable names

```{r message=FALSE, warning=FALSE}
colnames(df)
```

### 0.3 rename

```{r}
colnames(df) <- c("編號", "性別", "慢性病用藥狀況", "Comorbidity", "醫院代碼", "術前主述", "Biopsy_date_確診日期", "Biopsy_method", "Cell_Type", "多發性", "切片檢體腫瘤惡性度", "病理分期", "危險因子", "合併膀胱腫瘤", "合併CIS", "surgical_margin", "左右側", "腫瘤位置", "腫瘤大小", "完整檢體腫瘤惡性度", "pathological_stage", "術前水腎", "Lymphovascular_invasion", "Tumor_Necrosis", "有無針對UTUC化療", "針對UTUC化療型態", "化療處方", "NxUx_date", "NxUx_Access_method", "Bladder_cuff_resection", "other_bladder_cuff_method", "Lymphadenectomy_位置", "simultaneously_ipsilateral_adrenalectomy", "Post_operation_intravesical_CT_instillation", "Endoscopic_resection_date", "Endoscopic_Access_method", "Endoscopic_Energy_device", "endoscopic_ablation_次數", "post_ablation_Intra_cavitary_CT", "Salvage_Nephroureterectomy", "date_of_salvage_NU", "date_of_segmental_resection", "Segmental_resection_Access_method", "salvage_Nephroureterectomy", "Date_of_salvage_NU", "Clavien_Dindo_classification", "complication_list", "術後住院天數_天", "Residual_bladder_cuff", "Date_of_last_cystoscopy", "Bladder_UC_after_NUx_or_Endoscopic_or_Segmental_resection", "Date_of_Bladder_UC_recurrence_after_NUx_or_Endoscopic_or_Segmental_resection", "患側Upper_ureter_or_renal_pelvis_local_recurrence", "Date_of_Upper_ureter_or_renal_pelvis_recurrence", "患側Lower_ureter_or_bladder_cuff_local_recurrence", "Date_of_Lower_ureter_or_bladder_cuff_recurrence", "淋巴轉移及位置", "Date_of_LN_mets", "遠端轉移", "Date_of_distant_mets", "Disease_free", "Mortality", "Date_of_mortality", "長期_Complication", "手術到死亡間隔時間_月", "Post_OP_1_month_eGFR", "last_eGFR", "Lost_follow_up", "Longest_follow_up_month_for_BDFS_DFS", "Longest_follow_up_month_for_OS_CSS", "死亡檔最後追蹤日期", "死亡檔最長追蹤時間_月", "ECOG", "ASA", "身高", "體重", "生日", "診斷年紀", "術前Cr_level_mg_dl")
```

0.4 drop missing rows

```{r}
n_missing <- c()

for(i in 1:6345){
  if(sum(1*is.na(df[i,]))>50){
    n_missing <- c(n_missing, i)
  }
}


df1 <- df[-n_missing,]
```

****

## **1 Selection**

### 1.1 pckage

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

### 1.2 `select()`

```{r}
df2 <- df1%>%
  dplyr::select(編號, 性別, ECOG, 身高, 體重, 生日, Comorbidity, 腫瘤位置, 腫瘤大小, pathological_stage, Mortality, Date_of_mortality, 術前Cr_level_mg_dl, Post_OP_1_month_eGFR, 死亡檔最長追蹤時間_月)

df2
```

****

## **2 Type**

```{r}
df3 <- df2%>%
  transform(性別 = as.factor(性別))%>%
  transform(ECOG = as.factor(ECOG))%>%
  transform(身高 = as.numeric(身高))%>%
  transform(體重 = as.numeric(體重))%>%
  transform(生日 = as.Date(生日, "%m/%d/%Y"))%>%
  transform(Comorbidity = as.factor(Comorbidity))%>%
  transform(腫瘤位置 = as.factor(腫瘤位置))%>%
  transform(腫瘤大小 = as.factor(腫瘤大小))%>%
  transform(pathological_stage = as.factor(pathological_stage))%>%
  transform(Mortality = as.factor(Mortality))%>%
  transform(Date_of_mortality = as.Date(Date_of_mortality, "%m/%d/%Y"))%>%
  transform(術前Cr_level_mg_dl = as.numeric(術前Cr_level_mg_dl))%>%
  transform(Post_OP_1_month_eGFR = as.numeric(Post_OP_1_month_eGFR))%>%
  transform(死亡檔最長追蹤時間_月 = as.numeric(死亡檔最長追蹤時間_月))

df3
```

### 2.1 Factor

```{r}
df2%>%
  transform(性別 = as.factor(性別))%>%
  transform(ECOG = as.factor(ECOG))%>%
  transform(Comorbidity = as.factor(Comorbidity))%>%
  transform(腫瘤位置 = as.factor(腫瘤位置))%>%
  transform(腫瘤大小 = as.factor(腫瘤大小))%>%
  transform(pathological_stage = as.factor(pathological_stage))%>%
  transform(Mortality = as.factor(Mortality))%>%
  dplyr::select(性別, ECOG, Comorbidity, 腫瘤位置, pathological_stage, Mortality)
```

### 2.2 Numeric

```{r}
df2%>%
  transform(身高 = as.numeric(身高))%>%
  transform(體重 = as.numeric(體重))%>%
  transform(術前Cr_level_mg_dl = as.numeric(術前Cr_level_mg_dl))%>%
  transform(Post_OP_1_month_eGFR = as.numeric(Post_OP_1_month_eGFR))%>%
  transform(死亡檔最長追蹤時間_月 = as.numeric(死亡檔最長追蹤時間_月))%>%
  dplyr::select(身高, 體重, 術前Cr_level_mg_dl, Post_OP_1_month_eGFR, 死亡檔最長追蹤時間_月)
```

### 2.3 Date

```{r}
df2%>%
  transform(生日 = as.Date(生日, "%m/%d/%Y"))%>%
  transform(Date_of_mortality = as.Date(Date_of_mortality, "%m/%d/%Y"))%>%
  select(生日, Date_of_mortality)
```


****

### **2.4 Recap**{.tabset}

#### data frame

```{r}
df3
```

#### `summary()`

```{r}
summary(df3)
```

****

## **3 Detail** 

```{r}
df4 <- df3%>%
  mutate(BMI = factor(ifelse(is.na(身高)|is.na(體重),
                             NA,
                             ifelse(體重/(身高/100)^2<25,
                                    "0 Normal", "1 Overweight"))),
         .after = 體重)%>%
  dplyr::select(-身高, -體重)%>%
  mutate(Comorbidity = as.factor(
    ifelse(is.na(Comorbidity), NA,
                              ifelse(grepl("5 HTN", Comorbidity),
                                     "1 HTN",
                                     "0 No HTN"))
  ))%>%
  mutate(腫瘤位置 = as.factor(
    ifelse(is.na(腫瘤位置), "5 無法識別",
           ifelse(腫瘤位置=="0 non-visible", "0 non-visible",
                       ifelse(腫瘤位置=="1 腎盂", "1 腎盂",
                          ifelse(腫瘤位置=="2 上輸尿管"|
                                   腫瘤位置=="3 中輸尿管"|
                                   腫瘤位置=="4 下輸尿管"|
                                   腫瘤位置=="6 Urerter, unknown site", "2 輸尿管",
                                 ifelse(腫瘤位置=="5 膀胱袖口", "3 膀胱袖口",
                                        ifelse(腫瘤位置=="7 Not available", "5 無法識別",
                                               "4 多個位置"))))))
  ))%>%
  mutate(腫瘤大小 = as.factor(
    ifelse(is.na(腫瘤大小)|腫瘤大小=="Not available", "0 Not available",
           ifelse(grepl(3,腫瘤大小)|grepl(4,腫瘤大小), "2 >=2cm", "1 <2cm"))
  ))%>%
  mutate(pT = as.factor(
    ifelse(is.na(pathological_stage), NA,
                     ifelse(grepl("T0", pathological_stage)|
                              grepl("Tis", pathological_stage)|
                              grepl("Ta", pathological_stage), "1 T0/Tis/Ta",
                            ifelse(grepl("T1", pathological_stage), "2 T1",
                                   ifelse(grepl("T2", pathological_stage), "3 T2",
                                          ifelse(grepl("T3", pathological_stage), "4 T3",
                                                 ifelse(grepl("T4", pathological_stage), "5 T4", "0 None"))))))
  ),
         .after = pathological_stage)%>%
  mutate(pN = as.factor(
    ifelse(is.na(pathological_stage), NA,
                     ifelse(grepl("9 ", pathological_stage), "1 Nx",
                            ifelse(grepl("6 ", pathological_stage), "2 N0",
                                   ifelse(grepl("7 ", pathological_stage), "3 N1",
                                          ifelse(grepl("8 ", pathological_stage), "4 N2", "0 None")))))
  ),
         .after = pT)%>%
  mutate(pM = as.factor(
    ifelse(is.na(pathological_stage), NA,
                     ifelse(grepl("12 ", pathological_stage), "1 M0",
                            ifelse(grepl("11 ", pathological_stage), "2 M1", "0 None")))
  ),
         .after = pN)%>%
  dplyr::select(-pathological_stage)%>%
  mutate(Mortality = as.factor(
    ifelse(is.na(Mortality),
           NA,
           ifelse(grepl(0, Mortality),
                  "0 No",
                  ifelse(grepl(1, Mortality),
                         "1 UTUC related",
                         ifelse(grepl(2, Mortality),
                                "2 Non-UTUC related",
                                ifelse(grepl(3, Mortality),
                                       "3 死因不明",
                                       ifelse(grepl(4, Mortality),
                                              "4 Surgery related", NA))))))
  ))%>%
  mutate(Age_of_Mortality = ifelse(is.na(Date_of_mortality)|is.na(生日), NA,
                                   as.numeric(difftime(Date_of_mortality,
                                                       生日))),
         .after = Date_of_mortality)%>%
  mutate(Age_of_Mortality = round(Age_of_Mortality/365, 0))

df4
```

### 3.1 Functions

* `mutate()`
* `ifelse()`
  + `is.na()`
  + `grepl()`
* `round()`

****

### 3.2 Variables

****

#### 3.2.1 BMI

```{r}
df3%>%
  mutate(BMI = round(體重/(身高/100)^2, digits = 2),
         .after = 體重)%>%
  mutate(BMI.group = factor(ifelse(is.na(身高)|is.na(體重),
                             NA,
                             ifelse(體重/(身高/100)^2<25,
                                    "Normal", "Overweight"))),
         .after = BMI)%>%
  dplyr::select(身高, 體重, BMI, BMI.group)
```

#### 3.2.2 Comorbidity

```{r}
df3%>%
  mutate(Comorbidity.HTN = as.factor(
    ifelse(is.na(Comorbidity), NA,
                              ifelse(grepl("5 HTN", Comorbidity),
                                     "1 HTN",
                                     "0 No HTN"))
  ))%>%
  dplyr::select(Comorbidity, Comorbidity.HTN)
```


#### 3.2.3 腫瘤位置

```{r}
df3%>%
  mutate(腫瘤位置.group = as.factor(
    ifelse(is.na(腫瘤位置), "5 無法識別",
           ifelse(腫瘤位置=="0 non-visible", "0 non-visible",
                       ifelse(腫瘤位置=="1 腎盂", "1 腎盂",
                          ifelse(腫瘤位置=="2 上輸尿管"|
                                   腫瘤位置=="3 中輸尿管"|
                                   腫瘤位置=="4 下輸尿管"|
                                   腫瘤位置=="6 Urerter, unknown site", "2 輸尿管",
                                 ifelse(腫瘤位置=="5 膀胱袖口", "3 膀胱袖口",
                                        ifelse(腫瘤位置=="7 Not available", "5 無法識別",
                                               "4 多個位置"))))))
  ))%>%
  dplyr::select(腫瘤位置, 腫瘤位置.group)
```

#### 3.2.4 腫瘤大小

```{r}
df3%>%
  mutate(腫瘤大小.2cm = as.factor(
    ifelse(is.na(腫瘤大小)|腫瘤大小=="Not available", "0 Not available",
           ifelse(grepl(3,腫瘤大小)|grepl(4,腫瘤大小), "2 >=2cm", "1 <2cm"))
  ))%>%
  dplyr::select(腫瘤大小, 腫瘤大小.2cm)
```


#### 3.2.5 Pathological Stage

```{r}
df3%>%
  mutate(pT = as.factor(
    ifelse(is.na(pathological_stage), NA,
                     ifelse(grepl("T0", pathological_stage)|
                              grepl("Tis", pathological_stage)|
                              grepl("Ta", pathological_stage), "T0/Tis/Ta",
                            ifelse(grepl("T1", pathological_stage), "T1",
                                   ifelse(grepl("T2", pathological_stage), "T2",
                                          ifelse(grepl("T3", pathological_stage), "T3",
                                                 ifelse(grepl("T4", pathological_stage), "T4", "None"))))))
  ),
         .after = pathological_stage)%>%
  mutate(pN = as.factor(
    ifelse(is.na(pathological_stage), NA,
                     ifelse(grepl("9 ", pathological_stage), "Nx",
                            ifelse(grepl("6 ", pathological_stage), "N0",
                                   ifelse(grepl("7 ", pathological_stage), "N1",
                                          ifelse(grepl("8 ", pathological_stage), "N2", "None")))))
  ),
         .after = pT)%>%
  mutate(pM = as.factor(
    ifelse(is.na(pathological_stage), NA,
                     ifelse(grepl("12 ", pathological_stage), "M0",
                            ifelse(grepl("11 ", pathological_stage), "M1", "None")))
  ),
         .after = pN)%>%
  dplyr::select(pathological_stage, pT, pN, pM)
```

#### 3.2.6 Mortality

```{r}
df3%>%
  mutate(Mortality.na = as.factor(
    ifelse(is.na(Mortality),
           NA,
           ifelse(grepl(0, Mortality),
                  "0 No",
                  ifelse(grepl(1, Mortality),
                         "1 UTUC related",
                         ifelse(grepl(2, Mortality),
                                "2 Non-UTUC related",
                                ifelse(grepl(3, Mortality),
                                       "3 死因不明",
                                       ifelse(grepl(4, Mortality),
                                              "4 Surgery related", NA))))))
  ))%>%
  dplyr::select(Mortality, Mortality.na)
```

#### 3.2.7 Age of Mortality

```{r}
df3%>%
  mutate(Age_of_Mortality = ifelse(is.na(Date_of_mortality)|is.na(生日), NA,
                                   as.numeric(difftime(Date_of_mortality,
                                                       生日))),
         .after = Date_of_mortality)%>%
  mutate(Age_of_Mortality = round(Age_of_Mortality/365, 0))%>%
  dplyr::select(生日, Date_of_mortality, Age_of_Mortality)
```

## **4 Missing value**

```{r}
df5 <- df4%>%
  filter(if_all(everything(), ~ !is.na(.)))

df5
```

## **5 **

```{r}
df6 <- df5%>%
  select(編號, 性別, ECOG, BMI, Comorbidity, 腫瘤位置, 腫瘤大小, pT, pN, pM, Mortality, 死亡檔最長追蹤時間_月)%>%
  mutate(編號 = as.numeric(gsub(".*?([0-9]+).*", "\\1", 編號)))%>%
  mutate(性別 = as.numeric(gsub(".*?([0-9]+).*", "\\1", 性別)))%>%
  mutate(ECOG = as.numeric(gsub(".*?([0-9]+).*", "\\1", ECOG)))%>%
  mutate(BMI = as.numeric(gsub(".*?([0-9]+).*", "\\1", BMI)))%>%
  mutate(Comorbidity = as.numeric(gsub(".*?([0-9]+).*", "\\1", Comorbidity)))%>%
  mutate(腫瘤位置 = as.numeric(gsub(".*?([0-9]+).*", "\\1", 腫瘤位置)))%>%
  mutate(腫瘤大小 = as.numeric(gsub(".*?([0-9]+).*", "\\1", 腫瘤大小)))%>%
  mutate(pT = as.numeric(gsub(".*?([0-9]+).*", "\\1", pT)))%>%
  mutate(pN = as.numeric(gsub(".*?([0-9]+).*", "\\1", pN)))%>%
  mutate(pM = as.numeric(gsub(".*?([0-9]+).*", "\\1", pM)))%>%
  mutate(Mortality = as.numeric(gsub(".*?([0-9]+).*", "\\1", Mortality)))%>%
  mutate(死亡檔最長追蹤時間_月 = as.numeric(gsub(".*?([0-9]+).*", "\\1", 死亡檔最長追蹤時間_月)))

df6
```

## **6 Results**

### 6.1 data frame{.tabset}

#### table

```{r}
df5
```

#### `summary()`

```{r}
summary(df5)
```

### 6.2 numerical data frame{.tabset}

#### table

```{r}
df6
```

#### `summary()`

```{r}
summary(df6)
```